\documentclass[sigconf]{acmart}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{url}
\usepackage{pgfplots}
\usepackage{tabularx}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{pifont}

%LTeX: language=en-GB

%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% These commands are for a PROCEEDINGS abstract or paper.
\settopmatter{printacmref=true} % Removes citation information below abstract
%% Was false before, we need to check why its not working as intended
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in 

\acmConference[AEPRO 2026]{AEPRO 2026: Algorithm Engineering Projects}{March 16}{Jena, Germany}

% convert text to title case
% http://individed.com/code/to-title-case/

% that helps you to formulate your sentences
% https://www.deepl.com/translator

\pgfplotsset{compat=1.18}

\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
%% TODO: think about good title
\title{Image Text Enhancer: A Lightweight, Easy-to-Use, and Efficient C++ Library}
\subtitle{\large Algorithm Engineering 2026 Project}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.

\author{Daniel Motz}
\affiliation{%
  \institution{Friedrich Schiller University}
  \city{Jena}
  \country{Germany}}
\email{daniel.motz@uni-jena.de}

\author{Leonard Teschner}
\affiliation{%
  \institution{Friedrich Schiller University}
  \city{Jena}
  \country{Germany}}
\email{leonard.teschner@uni-jena.de}

\author{Mher Mnatsakanyan}
\affiliation{%
  \institution{Friedrich Schiller University}
  \city{Jena}
  \country{Germany}}
\email{mher.mnatsakanyan@uni-jena.de}

%% The abstract is a short summary of the work to be presented in the article.
\begin{abstract}
  % RQ Can a lightweight document enhancement pipeline be optimised to achieve strong multicore performance and practical WebAssembly deployment while preserving the enhancement behavior of the baseline implementation?
  TBC
\end{abstract}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{task parallelism, WebAssembly, OpenMP, image processing, C++}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
\label{sec:intro}
\begin{comment}
  Possible Research Questions
  \begin{enumerate}
    \item Easily extensible Open Source Pipeline, ready to use out of the box with default options giving a perfectly good image, completely written in C++ and licensed under MIT-0 and CeCILL-C, which are both commercially available (compared to e.g. unpaper) and has no big dependencies (unlike unpaper, which uses ffmpeg) and only uses CImg, which is portable (uses OpenMP), simple, self-contained, and lightweight.
    \item Evaluating Performance Gains: How do systematic code optimisations improve the processing speed of a baseline image enhancement tool?
    \item Scaling Performance in Image Processing: Identifying and mitigating inefficiencies in an initial software implementation for OCR 
  \end{enumerate}
\end{comment}
\textbf{RQ} Can a lightweight document enhancement pipeline be optimised to achieve strong multicore performance and practical WebAssembly deployment while preserving the enhancement behaviour of the baseline implementation? \\

\textbf{RQ1.} How much faster is the optimised pipeline than the baseline on native CPU execution? \\
\textbf{RQ2.} How well does the optimised pipeline scale under strong and runtime scaling from 1 to 8 cores? \\
\textbf{RQ3.} What runtime overhead does the WebAssembly version introduce compared to native execution, and is that overhead still acceptable for practical use?

\subsection{Background}
\label{sec:background}

The extraction of textual information from images --- from scanned books to even receipts captured by smartphone cameras --- have become an important part of everyday life \cite{alqudah_2015,review_degradation_image_enhancement_zhou_2023}. Any such extraction necessarily begins with the pre-processing of recorded images through techniques such as binarisation, noise removal, and geometric transformation, a process generally referred to as \textit{document enhancement}. Numerous tools focused on this niche have been proposed. Although these approaches feature broad capabilities, they are often associated with substantial complexity on the user's end and entail considerable software dependencies. These dependencies restrict applicability, particularly in settings where binary size is constrained, such as mobile and web platforms. This limitation assumes added importance in a context where most digital media interactions now take place on smartphones and application development is increasingly oriented toward edge-computing models. Consequently, efficient and platform-independent implementations that exploit parallelisation and other optimisation techniques are of considerable importance for reducing processing time and enhancing user satisfaction.

\subsection{Related Work}
\label{sec:related-work}

\begin{comment}
Since the introduction of WASM in 2017 \cite{WASM_Haas_2017}, research has progressed in a variety of ways. For instance, significant investment has gone into the 'internal' WASM runtime, either to add features or to investigate its performance and energy efficiency . Another area of research has investigated the use of WASM in various application \cite{wasm_runtimes_survey_zhang_2025} areas, such as web applications, the Internet of Things, and high-performance computing \cite{wasm_runtimes_survey_zhang_2025}. 
Of particular mention is the work of Chadha et al., who examine the use of WASM in the HPC domain in \cite{Mpi_Wasm_Chadha_2023}. Since parallelisation is essential for HPC applications, they have developed a WASM embedder that enables the MPI library to be used in WASM applications.  
\end{comment}
\begin{comment}
  \begin{enumerate}
    \item easy to use
    \item out of the box good results
    \item low dependencies (only CImg (doesnt have other dependencies) and OpenMP )
    \item only image enhancement, no OCR
    \item MIT-0 or CeCILL-C license (commercially usable)
    \item modular (users can choose which steps to apply)
  \end{enumerate}

  Tools die Document Enhancement bereits können:
  - OpenCV -> viel zu groß und abhöngig von vielen anderen Bibliotheken, BSD Lizenz
  - Tesseract - Open Source OCR Engine, die einige Bildverbesserungsfunktionen bietet, aber hauptsächlich auf OCR ausgerichtet ist, nicht einfach erweiterbar, preprocessing sind nicht als allgemeine Enhnancement Methoden gedacht, Apache lizenz
  - Leptonica - C-Bibliothek für Bildverarbeitung, hauptsächlich auf OCR ausgerichtet, BCD Lizenz
  - ImageMagick - Umfassende Kommandozeilen-Tool und Bibliothek für Bildbearbeitung, aber nicht speziell auf die Verbesserung von gescannten oder fotografierten Textbildern ausgerichtet, viele Abhängigkeiten, nicht so einfach erweiterbar, Apache Lizenz
  - Unpaper - spezialisiert auf Nachbearbeitung von Scans von Büchern, viele Abhängigkeiten (ffmpeg etc), nicht so einfach erweiterbar und stark spezialisiert, GPL Lizenz
\end{comment} 

Image enhancement and document preprocessing are covered by a wide range of open-source tools and commercial software. The scope ranges from general purpose image processing and OCR libraries to specialised tools for post-processing scanned documents.

ImageMagick \cite{imagemagick} is a suite for image processing that is available as a command line tool and as a library. It offers functions such as format conversion, resizing, filtering, and colour correction. However, ImageMagick is not specifically designed to enhance scanned or photographed images. Similar to OpenCV, the diverse features result in a large code base.

The C-based Leptonica library \cite{leptonica} is a general-purpose image processing and image analysis application. It is a dependency to both Tesseract and OpenCV. Although Leptonica is relatively lightweight compared to the other libraries, much like our approach, it primarily serves as a low-level library and does not offer ready-to-use, configurable processing pipelines.

\label{sec:specialised-post-processing}
A very similar tool to our approach is Unpaper \cite{unpaper}, which focuses on post-processing scanned documents. It offers features such as deskewing, binarisation, and noise reduction that have been developed specifically for this purpose. The tool also offers a CLI variant, with which pipeline steps can be configured and disabled. Unpaper requires the library ffmpeg, which in the standard configuration consumes >100MiB of space.

\begin{table}[t]
  \caption{Comparison of related tools.}
  \label{tab:tool-comparison-compact}
  \centering
  \begin{tabular}{lclc}
    \toprule
    \textbf{Tool} & \textbf{Lightweight} & \textbf{Licence} & \textbf{Commercial} \\
    \midrule
    ImageMagick & \ding{55}   & Own     & \ding{51} \\
    Leptonica   & \ding{51}   & Own     & \ding{51} \\
    Unpaper     & \ding{55}   & Mixed   & \ding{55} \\
    \bottomrule
  \end{tabular}
\end{table}

\

\label{sec:intro-optimisation}
\begin{comment}
  CoPilots Vorschlag:
  In addition to the tools and libraries described above, there is a large body of research on code optimisation and performance improvement in image processing applications. For example, various techniques have been proposed to optimise the performance of image processing algorithms, such as parallelisation, vectorization, and algorithmic improvements. However, these optimisations are often implemented in the context of specific libraries or applications and are not always easily transferable to other contexts. Moreover, the performance gains from these optimisations can vary widely depending on the specific use case and implementation details.
\end{comment}

\subsection{Our Contribution}
\label{sec:our-contribution}

% todo
% - State of the art binarisation etc
% - name, which algorithms we use

Unlike the existing tools and libraries described, our approach is designed to be minimalistic, extensible, and user-friendly. Instead of being a comprehensive image processing library or OCR engine, we focus on a lightweight, modular image enhancement pipeline, focused on printed and handwritten text, written entirely in C++ with only \href{https://www.cimg.eu/}{CImg} as a dependency, which is itself is lightweight has no further dependencies \cite{cimg}. Image Text Enhancer (ITE) has a lightweight build footprint, requiring only three build dependencies: OpenMP, X11, and JPEG. It can be imported as a library and also as a CLI executable---with 1.7 MiB on Linux.

Compared to OpenCV and ImageMagick, our work is characterised by its deliberate focus on document enhancement and compactness. Unlike Leptonica, our pipeline is specialised in the post-processing of scanned documents and unlike Unpaper, our binary including dependencies is $\approx$50$\times$ smaller.
In addition, our licensing model (MIT-0 and CeCILL-C) permits use in research and commercial applications. By combining minimal dependencies, a modular architecture, portability through C++ and OpenMP, and a commercially usable licence, our pipeline fills a gap between large-scale, general-purpose image processing libraries and highly specialised tools for document enhancement.

Furthermore, we systematically optimise and improve our baseline implementation and evaluate the performance gains through various experiments. In a proof of concept, we show that our approach can be executed on WASM platforms, with only minor performance drawbacks.

\subsection{Outline}
\label{sec:outline}

In section \ref{sec:pipeline}, we introduce the pipeline that we will implement and optimise. In section \ref{sec:optimisation}, we describe the optimisation techniques used. These are evaluated in section \ref{sec:experiments}, where we compare the performance of the pipeline under different settings to the baseline implementation. Finally, we provide an outlook on possible future work \ref{sec:future-work}. 

\section{Pipeline}
\label{sec:pipeline}
  
The pipeline is designed to be modular, enabling users to activate, deactivate and configure each stage as required, both through the CLI and the library. Some stages can intentionally not be disabled. These exceptions are marked in algorithm \ref{alg:enhance} with an \textbf{asterisk}. 

\begin{algorithm}
  \caption{Image Text Enhancer Pipeline}
  \label{alg:enhance}
  \begin{enumerate}
  \item Grayscale Conversion\textbf{*}
  \item Deskew
  \item Contrast Enhancement
  \item Denoise: (Adaptive) Gaussian and Median Blur
  \item Binarisation\textbf{*}: Otsu, Sauvola, Bataineh
  \item Morphological Operations
  \item Colour passthrough
\end{enumerate}
\end{algorithm}

The user can choose between different algorithms for certain stages---as these have different effects on the result, users can select the algorithm best suited to their application. For instance, Bataineh's adaptive binarisation method with automatic window adjustment \cite{adaptive_thresholding_methods_bataineh_2011} is helpful when binarising historical, handwritten documents with uneven writing and lighting, whereas Otsu's method \cite{Otsu1979} is a good choice for binarising typewritten documents with uniform font and lighting.
Standard configurations are provided for each step, enabling the pipeline to be used straight away. These configurations deliver good results for most use cases and are based on fixed recommendations from the authors. This means that the pipeline can be applied immediately. However, users can also customise the configurations to optimise the pipeline and improve their specific results.

We provide a command-line application that can be used independently of a program to access the pipeline. This enables the pipeline to be configured and used directly via the command line. 

\section{Optimisations}
\label{sec:optimisation}


\begin{comment}
  - Grayscale: parallel for, for pixel processing loop
  - Grayscale: pointer swapping instead of copying the image for each step -> no memory tests were conducted

  - Deskewing: replaced "rotate and project" method with a "radon transform". Instead of moving pixels, we extract foreground coordinates and mathematically project them
  - Deskewing: logic split: split detection and application. Angle is detected once and then applied to the gray and color image
  - Deskewing: total runtime reduced from ~95ms to ~7ms

  - Contrast Enhancement: parallel for unsing thread-local buffers to avoid false sharing/locking
  - Contrast Enhancement: precalculate the linear strecht formula and save in Lookup-Table. main loop performs O(1) lookup insted of floating point calculations
  - Contrast Enhancement: replacing if/else branching in pixel loop with std::min clamping, alowing the compiler to auto-vectorize using SIMD -> no assembly checks were done

  - Adaptive Gaussian Blur: removed std::vector allocations inside hot loop and using "ping-pong" strategie where image acted as destination buffer, swapping pointers at the end
  - Adaptive Median Filter: moving of buffer allocations outside of the pixel loop
  
  - CORE Utilites: returning pre-allocated buffers instead of heavy objects

  - Batainhe Binarisation: using integral images to calculate balck and red pixels instead of iterating over the complete window for each pixel -> reduced from windowsize * window size to 4 operations per pixel -> reduzierung von O(N \cdot W^2) zu O(N) 
  - Bataineh Binarisation: total runtime reduced from ~340ms to ~20ms
  - Sauvola Binarisation: branchless lookup
  - Suavola Binarisation: removing of the entire 8-bit image to "CImg<double>" twice. No uses the lightweigt fused vector builder

  - Morphological Operations: decomposing the 2d squre kernel operation intwo two 1d kernel operations (horizontal and vertical) -> reducing the complexity from O(K^2) to O(2K)
  - Morphological Operations: implementation of "sliding Window Min/max" algorithm using a monotonic queue -> calculation for any windowd size in amortized O(1) constant time
\end{comment}

We hypothesise that the baseline implementation fails to scale linearly with problem size, although such scaling would be desirable for image processing workloads. We therefore concentrated on identifying low-hanging fruits at the level of high-level algorithmic design, for example by selecting more efficient algorithms for specific tasks, while aiming to adhere to best practices such as avoiding false sharing and writing code amenable to auto-vectorisation. For the purposes of the paper, we focus on the stages, whose optimisations had the biggest effect. The full list of applied optimisations is documented in the project's repository. % todo: add Optimizations.txt to the git repository.

\subsection{Deskew}
Instead of using a "rotate-and-project" approach, which repeatedly rotates the image to find an appropriate angle, we applied the Radon transform, which gives an acceptable rotation after the first application. Also, we removed an overlook, where both the deskew for the  grayscale and colour image were calculated, if the colour passthrough option was enabled.  % todo: explain why Radon transform gives an acceptable solution if I want to

\subsection{Bataineh}
The main feature and bottleneck of the Bataineh algorithm is the adaptive sliding window, which takes into account the number of pixels from three categories---background, foreground and confusion \cite{adaptive_thresholding_methods_bataineh_2011}. This requires summing over areas of the image with masks reflecting these categories. The baseline approach was to sum using a nested for loop for each window. The optimised variant uses two integral image for background and confusion pixels, yielding a significant complexity reduction. We estimate our baseline implementation with $O(n \cdot w^2)$---where $n$ is the number of pixels and $w$ the window size---while the improved implementation is in $O(n)$.

\subsection{Grayscale}
% todo: unclear, because grayscale was already parallelised. Fix Optimizations.txt in this regard. Seemingly, only swapping was introduced, which yielded 40% time reduction... wth?

% todo: wieso sind die Filter besser geworden? Das hatte ich eigentlich schon so implementiert, was Mher jetzt schreibt, dachte ich.

\subsection{Morphology}
The pipeline only supports dilation and erosion, which can both be realised by linear filters. The baseline implementation used four nested for-loops\footnote{for colour images it is five, as it is done for each channel, but our approach is only used on binary images} for iterating the image and scanning the neighbourhood. Since the filter is separable, it is divided into a horizontal and vertical pass. Instead of scanning the neighbourhood, a queue is used to compute the min/max of a given 1D window in amortised $O(1)$.

% todo: why is the morphology worse then?
% todo: isn't the vertical pass really bad in regards to cache line use?

\subsection{Memory Management and Allocation}
\begin{comment}
  - Grayscale: pointer swapping instead of copying the image for each step -> no memory tests were conducted
  - Adaptive Gaussian Blur: removed std::vector allocations inside hot loop and using "ping-pong" strategie where image acted as destination buffer, swapping pointers at the end
  - Adaptive Median Filter: moving of buffer allocations outside of the pixel loop
  - Suavola Binarisation: removing of the entire 8-bit image to "CImg<double>" twice. No uses the lightweigt fused vector builder
  - CORE Utilites: returning pre-allocated buffers instead of heavy objects
\end{comment}

% todo: check these claims, e.g. wth do you mean by "pointer swap"?
To prevent increased memory consumption and long runtimes, the image is not copied at each step. Instead, the steps work in-place on the input image. For this purpose, a reference is passed through the pipeline. This approach is continued in grayscale conversion by replacing the final deep copy task of the image with a pointer swap of the original image pointer with the new one. 
Memory optimisations were made in further steps. For example, in the adaptive Gaussian blur filter and the adaptive median filter, the buffer allocations are removed from the pixel processing loop to reduce the number of memory allocations. Sauvola's binarisation algorithm uses a lightweight fused vector builder instead of the \textit{8-bit CImg<double>} image structure. Similarly, the core utilities return preallocated buffers instead of heavy objects.

\subsection{Algorithmic Complexity Reduction}
\begin{comment}
  - Batainhe Binarisation: using integral images to calculate balck and red pixels instead of iterating over the complete window for each pixel -> reduced from windowsize * window size to 4 operations per pixel -> reduzierung von O(N \cdot W^2) zu O(N) 
  - Morphological Operations: decomposing the 2d squre kernel operation intwo two 1d kernel operations (horizontal and vertical) -> reducing the complexity from O(K^2) to O(2K)
  - Morphological Operations: implementation of "sliding Window Min/max" algorithm using a monotonic queue -> calculation for any windowd size in amortized O(1) constant time
  - Deskewing: replaced "rotate and project" method with a "radon transform". Instead of moving pixels, we extract foreground coordinates and mathematically project them
\end{comment}
Several algorithms are implemented in the pipeline. Some of these are optimised through algorithmic improvements to reduce complexity. One example is Bataineh binarisation, which reduces the number of operations per pixel from $O(w \cdot n)$ to $O(n)$---where $w$ is the window and $n$ the image size---by using integral images: .


In morphological operations, the complexity is reduced from $O(n^2)$ to $O(2n)$---where $n$ is the number of pixels---by decomposing the 2D kernel operations into two 1D kernel operations for the horizontal and vertical dimensions. In addition, the sliding window min/max algorithm is implemented with a monotonic queue to enable computation for arbitrary window sizes in amortized $O(1)$ time. In the deskewing logic, the "rotate and project" method is replaced by a “Radon transform” method. Instead of moving the pixels, the coordinates of the foreground are extracted and mathematically projected.

% todo check the other O(1) claim, this does not sound right.

\subsection{Loop level Optimisations}
\begin{comment}
  - Contrast Enhancement: precalculate the linear strecht formula and save in Lookup-Table. main loop performs O(1) lookup insted of floating point calculations
  - Contrast Enhancement: replacing if/else branching in pixel loop with std::min clamping, alowing the compiler to auto-vectorize using SIMD -> no assembly checks were done
  - Sauvola Binarisation: branchless lookup
\end{comment}
To reduce the runtime of pixel processing loops, logical optimisations are made to the main loops that iterate over the pixels. In contrast enhancement, for example, the linear stretching formula is calculated in advance and stored in a lookup table. As a result, the main loop performs an $O(1)$ query per loop iteration instead of floating-point calculations. In addition, the if/else branch in the pixel processing loop is replaced by std::min clamping. This allows the compiler to auto-vectorize the loop with SIMD. In Sauvola binarisation, a branchless lookup is also implemented to improve performance.

\subsection{Architectial Refactoring}
\begin{comment}
  - Deskewing: logic split: split detection and application. Angle is detected once and then applied to the gray and color image
  - CORE Utilites: returning pre-allocated buffers instead of heavy objects
\end{comment}
In some cases, architectural changes were also made to improve performance. For example, in the deskewing logic, detection and application of the correction were separated. The angle is first detected and then applied to the grayscale and color images to avoid unnecessary calculations. In the core utilities, pre-built buffers are returned instead of creating heavy objects. This improves performance.

\section{Experiments}
\label{sec:experiments}

\subsection{Setup}
\label{sec:experimental-setup}
\begin{comment}
  Wir werden die performance (Laufzeit) der Pipeline anhand verschiedener Einstellungen testen. Bspw. unterschiediche Anzahl an Threads, verschieden große Bilder, verschiedene Kombinationen von Methoden.
  OS: WSL on Windows 11 (ist version relevant?)
  Hardware:
    i5 14600kf
    32gb ram
    affinity so wie im benchmark (0-7)
    desktop pc, ohne energiespaarmodus, ohne weitere apps im hintergrund
    auch lüfter an, damit keine temperatur throttles passieren könnten
hab nicht geschaut ob nur ausschließlich p-cores verwendet wurden, aber das wäre meine vermutung
\end{comment}

To evaluate the performance of the pipeline with the optimisations, various experiments are conducted and the results are compared with the baseline implementation. The experiments examine the runtime when processing images of different sizes with and without parallelisation.

The experiments assess three aspects: (1) single-threaded runtime versus image size, (2) eight-threaded runtime versus image size, and (3) strong-scaling behaviour with increasing core count for a large fixed image size. 

All stages of the pipeline are enabled, with Bataineh binarisation and the adaptive Gaussian and adaptive median filters selected.
% in den alten Benchmarks war adaptive gaussian, adaptive median und simple median an...
% todo: either redo benches without simple median or add it back here! It makes a lot of differrence in the strong scaling benchmark.

For our measurements, we use a machine with an AMD Ryzen 9950X3D 16-core processor (32 hyperthreads) running WSL2 with Ubuntu 24.04.1 LTS with 64 GB of RAM\footnote{It is clear that this comes with a performance penalty, however we can expect an even additive penalty across our conducted tests. See \underline{\href{https://www.phoronix.com/review/windows-11-wsl2-2025/2}{this}} blog post which tested the penalty on very similar hardware.}. Each core has a base frequency of 4.3 GHz, with boost frequency of around 5.5 GHz. Care has been taken to limit the number of background processes, so as not to affect performance.

We selected a small but representative set of document images spanning both document type and resolution. The benchmark includes textual documents of various sizes and degrees of detail (see Table \ref{tab:total-runtime} for details). These choices aim to capture common document-enhancement scenarios while also testing scalability to very large inputs.

% One test image is document text from the FUNSD dataset \cite{FUNSD_Jaume_2019}, which has been scanned and contains a lot of noise. Two additional test images are handwritten documents from the HTR ICFHR 2016 dataset \cite{HTR_Dataset_ICFHR_2016_Toselli}. Both of these images have a resolution of XxY and exhibit bleed-through, whereby text from other pages is visible. For colour documents, a test image from the LRDE Document Binarisation Dataset \cite{Efficient_Multiscale_Sauvola_Binarisation_Lazzara_Geraud_2013} is used. This is a scanned excerpt from a French magazine of very good quality containing many small sections of text. It has a resolution of XxY. A test image from the KITA imaging programme Forum is used to demonstrate a combination of historical text, UI elements and colour. All images are provided as JPG files.

% To test performance with larger images, the KITA image, which has an initial resolution of XxY, is upscaled to 2k, 4k, 8k and 16k using the ImageUpscaler tool \cite{imageupscaler}. All images are processed using the pipeline and the runtime is measured for each step. Additionally, the images are processed using one and eight threads to evaluate the effects of parallelisation.

\subsection{Results}
\label{sec:results}

Figures \ref{fig:opt-vs-notopt-8cores} and \ref{fig:opt-vs-notopt-1core} show the development of runtime with increasing image resolution for the baseline and optimised implementations. Unsurprisingly, the runtime $t$ increases with the problem size $s$. We found, that the optimised implementation scales almost linear, surpassing the baseline. For $c_1, c_2 < 1.7 \cdot 10^{-5}$ we observe

\begin{align*}
  t_{\text{baseline}} &\propto c_1 \cdot s^{1.660} \\
  t_{\text{optimised}} &\propto c_2 \cdot s^{1.071}
\end{align*}


% multithreaded runtime scaling
\begin{figure}[htbp]
\centering
\begin{tikzpicture}
  \begin{axis}[
    xmode=log,
    ymode=log,
    title={Multithreaded runtime scaling (8 Cores)},
    xlabel={Number of pixels (Megapixel)},
    ylabel={Runtime (ms)},
    xmin=0.5, xmax=140,
    ymin=20,
    grid=major,
    legend pos=south east,
    width=\columnwidth,
    height=6cm,
    table/col sep=comma,
  ]

  % Baseline implementation
  \addplot[
    color=blue,
    mark=square,
    thick,
    point meta=explicit,
    visualization depends on={\thisrow{trials} \as \trialcount},
    nodes near coords={
      \pgfmathparse{\trialcount==1000 ? "" : int(\trialcount)}
      \ifx\pgfmathresult\empty\else n=\pgfmathresult\fi
    },
    every node near coord/.append style={
      font=\scriptsize,
      text=gray!70,
      fill=white,
      fill opacity=0.8,
      text opacity=1,
      inner sep=1pt,
      rounded corners=1pt,
      anchor=south west,
      xshift=2pt,
      yshift=2pt
    }
  ] table[
    x expr=\thisrow{total_pixels}/1000000,
    y=avg_ms,
    meta=trials
  ] {graphics/bench_8_main.csv};
  \addlegendentry{baseline}

  % Optimized implementation
  \addplot[
    color=red,
    mark=x,
    thick,
    point meta=explicit,
    visualization depends on={\thisrow{trials} \as \trialcount},
    nodes near coords={
      \pgfmathparse{\trialcount==1000 ? "" : int(\trialcount)}
      \ifx\pgfmathresult\empty\else n=\pgfmathresult\fi
    },
    every node near coord/.append style={
      font=\scriptsize,
      text=gray!70,
      fill=white,
      fill opacity=0.8,
      text opacity=1,
      inner sep=1pt,
      rounded corners=1pt,
      anchor=north east,
      xshift=-2pt,
      yshift=-2pt
    }
  ] table[
    x expr=\thisrow{total_pixels}/1000000,
    y=avg_ms,
    meta=trials
  ] {graphics/bench_8_optimized.csv};
  \addlegendentry{optimised}

  \end{axis}
\end{tikzpicture}

\caption{Runtime as a function of image size for the baseline and optimised implementations on 8 cores. Each plotted point corresponds to the mean runtime over repeated measurements for the respective input image. Labels are shown only for measurements whose sample count differs from the standard case of 1000 repetitions. For clarity, all but one documents with image size of $\approx 10^1$ Megapixels were omitted, as the results were almost identical.}
\Description{Log-log plot of runtime versus image size in megapixels for baseline and optimised implementations running on 8 CPU cores. Each point shows the mean runtime for one input image over repeated measurements. Runtime increases with image size for both methods, while the optimised implementation remains consistently faster across the full range. Notice, that there is no measurement for the 16K image in the baseline implementation.}
\label{fig:opt-vs-notopt-8cores}
\end{figure}

% single threaded runtime scaling
\begin{figure}[htbp]
\centering
\begin{tikzpicture}
  \begin{axis}[
    xmode=log,
    ymode=log,
    title={Single-threaded runtime scaling},
    xlabel={Number of pixels (Megapixel)},
    ylabel={Runtime (ms)},
    xmin=0.5, xmax=140,
    ymin=20,
    grid=major,
    legend pos=south east,
    width=\columnwidth,
    height=6cm,
    table/col sep=comma,
  ]

  % Baseline implementation
  \addplot[
    color=blue,
    mark=square,
    thick,
    point meta=explicit,
    visualization depends on={\thisrow{trials} \as \trialcount},
    nodes near coords={
      \pgfmathparse{\trialcount==1000 ? "" : int(\trialcount)}
      \ifx\pgfmathresult\empty\else n=\pgfmathresult\fi
    },
    every node near coord/.append style={
      font=\scriptsize,
      text=gray!70,
      fill=white,
      fill opacity=0.8,
      text opacity=1,
      inner sep=1pt,
      rounded corners=1pt,
      anchor=south west,
      xshift=2pt,
      yshift=2pt
    }
  ] table[
    x expr=\thisrow{total_pixels}/1000000,
    y=avg_ms,
    meta=trials
  ] {graphics/bench_1_main.csv};
  \addlegendentry{baseline}

  % Optimized implementation
  \addplot[
    color=red,
    mark=x,
    thick,
    point meta=explicit,
    visualization depends on={\thisrow{trials} \as \trialcount},
    nodes near coords={
      \pgfmathparse{\trialcount==1000 ? "" : int(\trialcount)}
      \ifx\pgfmathresult\empty\else n=\pgfmathresult\fi
    },
    every node near coord/.append style={
      font=\scriptsize,
      text=gray!70,
      fill=white,
      fill opacity=0.8,
      text opacity=1,
      inner sep=1pt,
      rounded corners=1pt,
      anchor=north east,
      xshift=-2pt,
      yshift=-2pt
    }
  ] table[
    x expr=\thisrow{total_pixels}/1000000,
    y=avg_ms,
    meta=trials
  ] {graphics/bench_1_optimized.csv};
  \addlegendentry{optimised}

  \end{axis}
\end{tikzpicture}

\caption{}
\Description{Log-log plot of runtime versus image size in megapixels for baseline and optimised implementations running on 8 CPU cores. Each point shows the mean runtime for one input image over repeated measurements. Runtime increases with image size for both methods, while the optimised implementation remains consistently faster across the full range. Notice, that there is no measurement for the 16K image in the baseline implementation.}
\label{fig:opt-vs-notopt-1core}
\end{figure}

% strong scaling
\begin{figure}[htbp]
\centering

Strong scaling

\begin{minipage}{0.48\columnwidth}
\centering
\begin{tikzpicture}
  \begin{axis}[
    xmode=log,
    ymode=log,
    log basis x=2,
    log basis y=2,
    title={baseline},
    xlabel={Number of threads},
    ylabel={Speedup},
    xmin=1, xmax=32,
    ymin=1, ymax=32,
    xtick={1,2,4,8,16,32},
    grid=major,
    width=\linewidth,
    height=6cm,
    table/col sep=comma,
    legend to name=strongscalinglegend,
    legend columns=3,
  ]

  % 2K image
  \addplot[
    color=orange,
    mark=asterisk,
    thick
  ] table[
    x=threads,
    y=speedup
  ] {graphics/bench_strong_main_2k.csv};
  \addlegendentry{2K image}

  % 16K image
  \addplot[
    color=purple,
    mark=diamond,
    thick
  ] table[
    x=threads,
    y=speedup
  ] {graphics/bench_strong_optimized_2k.csv};
  \addlegendentry{16K image}

  % Ideal scaling
  \addplot[
    color=black,
    densely dashed,
    thick
  ] coordinates {
    (1,1) (2,2) (4,4) (6,6) (8,8) (10,10)
    (12,12) (14,14) (16,16) (32,32)
  };
  \addlegendentry{ideal}

  \end{axis}
\end{tikzpicture}
\end{minipage}
\hfill
\begin{minipage}{0.48\columnwidth}
\centering
\begin{tikzpicture}
  \begin{axis}[
    xmode=log,
    ymode=log,
    log basis x=2,
    log basis y=2,
    title={optimised},
    xlabel={Number of threads},
    ylabel={Speedup},
    xmin=1, xmax=32,
    ymin=1, ymax=32,
    xtick={1,2,4,8,16,32},
    grid=major,
    width=\linewidth,
    height=6cm,
    table/col sep=comma,
  ]

  % 2K image
  \addplot[
    color=orange,
    mark=asterisk,
    thick
  ] table[
    x=threads,
    y=speedup
  ] {graphics/bench_strong_optimized_2k.csv};

  % 16K image
  \addplot[
    color=purple,
    mark=diamond,
    thick
  ] table[
    x=threads,
    y=speedup
  ] {graphics/bench_strong_optimized_16k.csv};

  % Ideal scaling
  \addplot[
    color=black,
    densely dashed,
    thick
  ] coordinates {
    (1,1) (2,2) (4,4) (6,6) (8,8) (10,10)
    (12,12) (14,14) (16,16) (32,32)
  };

  \end{axis}
\end{tikzpicture}
\end{minipage}

\vspace{0.4em}
\pgfplotslegendfromname{strongscalinglegend}

\caption{Strong scaling behaviour of the baseline and optimised implementations. Speedup is shown as a function of thread count for two input images (2K and 16K). The dashed black line indicates ideal linear scaling.}
\Description{Two side-by-side log-log plots of speedup versus number of threads. The left panel shows the baseline implementation and the right panel shows the optimised implementation. Each panel contains curves for a 2K image and a 16K image, plus a dashed ideal-scaling reference line. A single shared legend is shown below the plots.}
\label{fig:strong-scaling-main-vs-optimised}
\end{figure}

The results for runtime scaling are seemingly good. We observe, that the optimisations lead to a significant improvement of the runtime, which is consistent with growing input size.


The results of the individual phases show a significant reduction in runtime in the deskewing and Bataineh binarisation phases when comparing the baseline and optimised implementations. Table \ref{tab:pahse-runtime} shows the runtime of the individual phases for the KITA test image with a resolution of 3840x2254 on 8 cores. It can be seen that the optimisations lead to a massive reduction in phase runtime of over 98\% for Bataineh binarisation in particular. The deskewing phase also benefits from the optimisations and shows a reduction in phase runtime of over 98\%. In other phases, such as grayscale and the filter, a significant reduction in phase runtime can also be observed, while in the morphological ops phase, a slight increase in phase runtime can be seen. WHY?????. Overall, this leads to a reduction in the total runtime of over 95\%. In absolute terms, the runtime improvement is mainly due to the optimisation of Bataineh binarisation, as this phase accounts for the largest share of the total runtime in the baseline implementation.

% Tabelle mit den Gesamtlaufzeiten fuer die verschiedenen Bilder und Einstellungen
\begin{table}[htbp]
\caption{Comparison of total runtimes (8 Cores)}
\label{tab:total-runtime}
\footnotesize
\begin{tabularx}{\columnwidth}{l r r r r}
\toprule
\textbf{Testbild} & \textbf{resolution} & \textbf{baseline (ms)} & \textbf{optimised (ms)} & \textbf{Speedup} \\
\midrule
Text Small 1K    & 1017x597 & 130,577  & 42,511 & 3,07$\times$ \\
Text 2K          & 2048x1202 & 1.770,138 & 188,975  & 9,37$\times$ \\
Magazine 2K      & 2516x3272 & 16.701,170 & 649,924 & 25,70$\times$ \\
Text 4K          & 3840x2254 & 16.328,084 & 781,634 & 20,89$\times$ \\
Text 8K          & 7680x4508 & 114.607,067 & 3.015,428 & 38,01$\times$ \\
Text 16K         & 15360x9017 & -\textsuperscript{*} & 15.181,484 & -\textsuperscript{*} \\
\bottomrule
\multicolumn{4}{l}{\scriptsize \textsuperscript{*}Could not be calculated due to massive runtime}
\end{tabularx}
\end{table}


\begin{table}[htbp]
\caption{Comparison of phase runtimes (3840x2254, 8 Cores)}
\label{tab:pahse-runtime}
\small
\centering
\begin{tabular}{l rr r}
\toprule
\textbf{Phase} & \textbf{baseline (ms)} & \textbf{optimised (ms)} & \textbf{diff (\%)} \\
\midrule
Init \& Copy    & 65,458  & 63,672  & -2,728\% \\
Grayscale       & 30,950  & 17,753  & -42,640\% \\
\textbf{Deskew} & \textbf{159,405} & \textbf{3,062} & \textbf{-98,079\%} \\
Filters (Sum)   & 483,312 & 380,467 & -21,279\% \\
\textbf{Bataineh} & \textbf{15.507,105} & \textbf{217,347} & \textbf{-98,598\%} \\
Morphological Ops & 66,663 & 93,05 & +39,583\% \\
Color Preservation & 4,129 & 4,089 & -0,969\% \\
\textbf{Total} & \textbf{16.328,084} & \textbf{781,634} & \textbf{-95,213\%} \\
\bottomrule
\end{tabular}
\end{table}

% todo: why does init and copy take so much time?

Table \ref{tab:breakdown} shows the detailed phase runtime for the different phases and images. It can be seen that the optimisations lead to a significant reduction in runtime in all phases, especially in Bataineh binarisation, which accounts for the largest share of the total runtime in the baseline implementation. For larger images such as the 8K text image, it can be seen that the image becomes too large for the system, as there is a sharp increase in the Init \& Copy phase, which indicates a memory limitation. 

\begin{table*}[t]
\caption{Detailed runtimes for certain pipeline stages}
\label{tab:breakdown}
\small
\centering
\begin{tabular}{l c c c c c c}
\toprule
\textbf{Phase} & \textbf{1K Small} & \textbf{2K Text} & \textbf{2K Mag.} & \textbf{4K Text} & \textbf{8K Text} \\
\midrule
Init \& Copy & 1,29  & 4,20  & 61,83  & 65,46  & 241,93  \\
Grayscale & 0,60  & 2,95  & 27,68  & 30,95  & 92,45  \\
Deskew & 24,22  & 59,89  & 28,90  & 159,41  & 518,54  \\
Binarisation & 71,00  & 1.555,57  & 15.999,83  & 15.507,11  & 111.495,99 \\
Filter (Sum) & 28,74 & 125,14 & 520,73 & 483,31 & 1.968,43 \\
\midrule
\textbf{TOTAL} & \textbf{130,58} & \textbf{1.770,14} & \textbf{16.701,17} & \textbf{16.328,08} & \textbf{114.607,07} \\
\bottomrule
\end{tabular}
\end{table*}

\section{Future Work}
\label{sec:future-work}

\begin{enumerate}
  \item Use heuristics for selecting algorithms, e.g. binarisation
  \item better deskew
  \item runtime experiments on WASM
  \item try to optimise for WASM (reduce the performance gap)
  \item What do we do to avoid false sharing?
  \item Look into blocking code
\end{enumerate}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{literature}


\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.